A Classical Density-Functional Theory for Describing Water Interfaces

### 2011-09-06 paper/project.tex
Fixed entropy in rods, edits to paper, adding xassoc for single rod

\subsection{Ideal gas functional}
The first term is the ideal gas free energy functional,
which incorporates effects of kinetic energy.  The ideal gas
functional is given by
\begin{equation}\label{idealgas}
  F_{id}[n] = k_B T \int n(\xx)\left( \ln{n(\xx)} - 1\right) d\xx
\end{equation}
where n(\xx) is the density of water molecules.  The ideal gas free
energy functional has the property that the contact density at a hard
surface $n_c$ is proportional to the pressure on that surface,
according to the contact value theorem
\begin{equation}\label{contactvaluethm}
  p = n_c k_BT \:,
\end{equation}
which also leads to the property that the free energy of solvation of
a hard solute is proportional to its volume in the limit of small
solutes.  These properties are retained by the total functional,
provided the remaining terms are \emph{purely
  nonlocal}~\cite{ashcroft?}\needsworknow{I can't find this citation anywhere!}.

@@

Figure \ref{fig:energy-vs-diameter} shows the free energy per area versus
radius for the range of rods studied, with radii from 0.05 nm to 1.5 nm. After 
about 0.5 nm or so, this reaches
an asymptote equal to the surface tension. From the inset in Figure 
\ref{fig:surface-tension}, it is clear that the surface tension at room 
temperature is approximately 72 mN/m. This is very close to the value from
the larger diameter single rod energy per area, which is about 75~mN/m. 
\needsworknow{Here it will be beneficial to add some discussion of any
relevant experimental work or other simulations that deal with one rod-like 
thing in water. Do we have anything currently matching this?}

@@

\bibliography{paper}% Produces the bibliography via BibTeX.

>>>>

\subsection{Ideal gas functional}
The first term is the ideal gas free energy functional,
which incorporates effects of kinetic energy.  The ideal gas
functional is given by
\begin{equation}\label{idealgas}
  F_{id}[n] = k_B T \int n(\xx)\left( \ln{n(\xx)} - 1\right) d\xx
\end{equation}
where n(\xx) is the density of water molecules.  The ideal gas free
energy functional has the property that the contact density at a hard
surface $n_c$ is proportional to the pressure on that surface,
according to the contact value theorem
\begin{equation}\label{contactvaluethm}
  p = n_c k_BT \:,
\end{equation}
which also leads to the property that the free energy of solvation of
a hard solute is proportional to its volume in the limit of small
solutes.  These properties are retained by the total functional,
provided the remaining terms are \emph{purely nonlocal}. This requires
that
\begin{equation}
 \frac{\delta^2F_{\textit{excess}}[n]}{\delta n(\xx_1)~\delta n(\xx_1)}\neq\infty
\end{equation}
where $F_{\textit{excess}}[n]$ describes all the remaining terms in the functional.
This ensures that any higher order terms in the Taylor expansion
of $F[n+\Delta n]$ are very small.

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/xassoc-single-rod}
\end{center}
\caption{ Number of bonds per molecule near the interface with a single
hydrophobic rod. Compare with Walther\cite{walther2001carbon} fig 9. }
\label{fig:xassoc-single-rod}
\end{figure}

Figure \ref{fig:energy-vs-diameter} shows the free energy per area versus
radius for the range of rods studied, with radii from 0.05 nm to 1.5 nm. After 
about 0.5 nm or so, this reaches
an asymptote equal to the surface tension. From the inset in Figure 
\ref{fig:surface-tension}, it is clear that the surface tension at room 
temperature is approximately 72 mN/m. This is very close to the value from
the larger diameter single rod energy per area, which is about 75~mN/m. 
\needsworknow{Discussion of Walther 2001 paper\cite{walther2001carbon}}

@@

\bibliographystyle{unsrt}
\bibliography{paper}% Produces the bibliography via BibTeX.
